---
title: "Week1"
author: "Erika Lee"
date: "2023-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
source("setup.R")
```

```{r}
library(palmerpenguins)
library(dplyr)
data(penguins)
```

### Week 1 Assignment:

Data Types

```{r}
class(penguins)
str(penguins)
```

```{r}
class(penguins$species)
str(penguins$species)
```

Creating Lists

```{r}
myList <- list("apple", 1993, FALSE, penguins)
str(myList)
```

```{r}
names(myList) <- c("fruit", "year", "logic", "data")
names(myList)
```

Indexing

```{r}
# for lists we use double brackes [[]]
myList[[1]]

myList[["data"]]
```

```{r}
#for indexing vectors
# for vectors we use single brackets []
myVector <- c("apple", "banana", "pear")
myVector[2]
```

```{r}
# indexing data frames
# dataframe[row(s), columns()]
penguins[1:5, 2]

penguins[1:5, "island"]

penguins[1, 1:5]

penguins[1:5, c("species","sex")]

penguins[penguins$sex=='female',]

# $ for a single column
penguins$species
```

Dplyr Package:

Subsetting rows with filter()

```{r}
filter(penguins, species == "Adelie")

filter(penguins, species != "Adelie")

filter(penguins, island %in% c("Dream", "Torgersen") & !is.na(bill_length_mm))
```

Selecting columns with select()

```{r}
# Select two specific variables
select(penguins, species, sex)

# Select a range of variables
select(penguins, species:flipper_length_mm)

# Rename columns within select
select(penguins, genus = species, island)

# Select column variables that are recorded in mm
select(penguins, contains("mm"))
```

Creating new variables with mutate

```{r}
# New variable that calculates bill length in cm
mutate(penguins, bill_length_cm = bill_length_mm/10)

# mutate based on conditional statements
mutate(penguins, species_sex = if_else(sex == 'male', paste0(species,"_m"), paste0(species, "_f")))
```

Piping

```{r}
df1 <- filter(penguins, island == "Dream")
df2 <- mutate(df1, flipper_length_cm = flipper_length_mm/10)
df3 <- select(df2, species, year, flipper_length_cm)

print(df3)
```

```{r}
penguins %>% 
  filter(island == "Dream") %>% 
  mutate(flipper_length_cm = flipper_length_mm/10) %>% 
  select(species, year, flipper_length_cm)
```

Summarize data by a specific group (e.g. average body mass)

```{r}
penguins %>% 
  group_by(species) %>% 
  summarise(body_mass_avg = mean(body_mass_g, na.rm = TRUE))
```

Get a count of something

```{r}
penguins %>% 
  group_by(species, year) %>% 
  summarise(n_observations = n())
```

7.  Visualize

    ```{r}
    #install.packages("ggplot2")
    library(ggplot2)
    ```

General structure for ggplot:

ggplot(data = \<DATA\>) +

\<GEOM_FUNCTION\>(mapping = aes(\<MAPPINGS\>))

Histograms:

```{r}
# Histogram example: flipper length by species
ggplot(penguins) +
  geom_histogram(aes(x = flipper_length_mm, fill = species), alpha = 0.5, position = "identity") +
  scale_fill_manual(values = c("darkorange","darkorchid","cyan4"))
```

Bar plot:

```{r}
ggplot(penguins) +
  geom_bar(mapping = aes(x = island, fill = species))
```

Customized simple bar plot:

```{r}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"), 
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Scatterplots:

```{r}
ggplot(penguins) +
  geom_point(mapping = aes(x = body_mass_g, y = flipper_length_mm, color = species))
```

8.  Read and write data

There are many ways to read and write data in R. To read in .csv files, you can use `read_csv()` which is included in the Tidyverse with the `readr` package, and to save csv files use `write_csv()`. The `readxl` package is great for reading in excel files, however it is not included in the Tidyverse and will need to be loaded separately.

### Week 1 Notes

2023.10.18

**Git command review:**

Order of commands to create repository

1.  git clone { https code} - only use once on a computer (at beginning of project)

2.  git status - tells you the status of git files - can use this whenever!

3.  Step 1. git pull - pulls any edits made somewhere else

    -   prevents any issues with conflicting files

    1.  Creating a readme file - the homepage for a repository
        -   go to repository on github and see if there is already a README.file - if you set up repository with the readme file clicked, then it will already be there.
        -   If README was not created.... in r - go to file \> new file \> new markdown file
        -   name as README.md and save
            -   .md documents can be treated almost exactly as rmarkdown files... can use visual editor with .md, change header level, insert images, etc.
        -   push this to repository

4.  Step 2. git add {insert name of files to be added} - will add everything that is red or unchecked, OR you can just check them in the git tab on r

5.  Step 3. git commit -m "message in quotes"

    1.  Can go to check boxes in git tab and type commit message in git GUI window OR
    2.  add git commit -m "..." to terminal...

6.  Step 4. git push

    1.  Can push with green up arrow in git tab OR
    2.  Type git push into terminal

**Git Set Up Details:**

Git Username Set up - complete

```{r}
## install if needed (do this exactly once):
install.packages("usethis")

library(usethis)
use_git_config(user.name = "Erika Lee", user.email = "erika.lee@colostate.edu")
```

Git token, only need to do this once - complete

```{r}
gitcreds::gitcreds_set()
```
